# Practical Machine Learning Assignment

##Olumide 

February 5,2019

Introduction:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

### Required Packages
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(RGtk2)
library(rattle)
library(randomForest)
library(dplyr)
library(lubridate)
library(corrplot)
library(e1071)
```
### Download and pull up Dataset
```{r}
data.train<- read.csv("pml-training.csv")
data.test<- read.csv("pml-testing.csv")
```
```{r}
dim(data.train)

```
### Convert date and add new variable (Day)

```{r}
data.train$cvtd_timestamp<- as.Date(data.train$cvtd_timestamp, format = "%m/%d/%Y %H:%M")
data.train$Day<-factor(weekdays(data.train$cvtd_timestamp))
```
### Explore Data

```{r}
table(data.train$classe)
```
```{r}
prop.table(table(data.train$classe)) 
```
```{r}
prop.table(table(data.train$user_name))
```
```{r}
prop.table(table(data.train$user_name,data.train$classe),1)  
```
```{r}
prop.table(table(data.train$user_name,data.train$classe),2) 
```
```{r}
prop.table(table(data.train$classe, data.train$Day),1) 
```
```{r}
qplot(x=Day, fill=classe, data = data.train)
```

#### Remove columns with NA missing values
```{r}
data.train <- data.train[, colSums(is.na(data.train)) == 0]
data.test <- data.test[, colSums(is.na(data.test)) == 0] 
```
#### Remove columns that are not relevant to accelerometer measurements.
```{r}
classe<- data.train$classe
trainRemove<- grepl("^X|timestamp|window", names(data.train))
data.train<- data.train[, !trainRemove]
trainCleaned<- data.train[, sapply(data.train, is.numeric)]
trainCleaned$classe<- classe
testRemove<- grepl("^X|timestamp|window", names(data.test))
data.test<- data.test[, !testRemove]
testCleaned<- data.test[, sapply(data.test, is.numeric)]
```
### check for data dimensions
```{r}
dim(trainCleaned)
```
```{r}
dim(testCleaned)
```


### Create Train and Test data sets:
```{r}
set.seed(244534)
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```
###Data Modelling:
#####Indetifying significant variables:
#####We will fit a predictive model using Random Forest algorithm as it gives important variables and removes multicollinearity  and outliers. We will also use 5-fold cross validation when applying the algorithm.
```{r}
controlRf <- trainControl(method="cv", 5)
rfmod<- train(classe ~., data=trainData, method="rf", trControl=controlRf, importance=TRUE, ntree=100)
```
```{r}
rfmod
```
###Accuracy of the model on Validated data set:
```{r}
predictRfmod<- predict(rfmod, testData)
confusionMatrix(testData$classe, predictRfmod)
```
```{r}
accuracy <- postResample(predictRfmod, testData$classe)
accuracy
```
```{r}
Error <- 1 - as.numeric(confusionMatrix(testData$classe, predictRfmod)$overall[1])
Error
```
### So, the estimated accuracy of the model is 99.3373% and the estimated out-of-sample error is 0.66%.

###Predicting on Test Data Set
```{r}
myresult <- predict(rfmod, testCleaned[, -length(names(testCleaned))])
myresult
```
###Correlation Matrix
```{r}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="circle")
```
###Tree Visualization
```{r}
rtree<- rpart(classe ~ ., data=trainData, method="class")
prp(rtree)
```